Metric Spaces with Expensive Distances

### 2019-04-14 paper/draft_esa.tex
interim

\title{Metric Spaces with Expensive Distances}
\titlerunning{Metric Spaces with Expensive Distances}
% uncomment for LIPICS 
\author{Michael Kerber}{TU Graz}{kerber@tugraz.at}{https://orcid.org/0000-0002-8030-9299}{Supported by Austrian Science Fund (FWF) grant number P 29984-N35}
\author{Arnur Nigmetov}{TU Graz}{nigmetov@tugraz.at}{}{}
\authorrunning{M. Kerber and A. Nigmetov}

@@

\Copyright{Michael Kerber and Arnur Nigmetov}
% END uncomment for LIPICS

@@

\begin{figure}[h]
\centering
\includegraphics[width=6cm]{intro_example.eps}
\caption{The compute distances are shown as edges in a graph. Note that the exact distance
of $p_1$ and $p_2$ is unknown. The shortest path from $p_1$ to $p_2$ has length $9$, which clearly
constitutes an upper bound on the distance by triangle inequality.
However, we can also infer that $\dist(p_1,p_2)\geq 3$:
otherwise, the path from $p_3$ to $p_4$ via $p_1$ and $p_2$
would be shorter than the edge $(p_3,p_4)$, again contradicting
triangle inequality.}
\label{fig:1st_example}
\end{figure}

@@

    \addplot+ table [x=n_points, y=edges_quasi_sorted_greedy, col sep=comma] {all_methods_dim_2_eps_0.1.txt};
    \addlegendentry{Blind quasi-sorted greedy}

    \addplot+ table [x=n_points, y=edges_quasi_sorted_shaker, col sep=comma] {all_methods_dim_2_eps_0.1.txt};
    \addlegendentry{Blind quasi-sorted shaker}

    \addplot+ table [x=n_points, y=edges_wspd, col sep=comma] {all_methods_dim_2_eps_0.1.txt};
    \addlegendentry{WSPD}

    \addplot+ table [x=n_points, y=edges_blind_random, col sep=comma] {all_methods_dim_2_eps_0.1.txt};
    \addlegendentry{Blind random}

    \addplot+ table [x=n_points, y=edges_blind_random_lower_bound_first, col sep=comma] {all_methods_dim_2_eps_0.1.txt};
    \addlegendentry{Blind random, lower bound first}

    \addplot+ table [x=n_points, y=edges_blind_greedy, col sep=comma] {all_methods_dim_2_eps_0.1.txt};
    \addlegendentry{Blind greedy}

    \addplot+ table [x=n_points, y=edges_greedy, col sep=comma] {all_methods_dim_2_eps_0.1.txt};
    \addlegendentry{Greedy}

@@

    \addplot+ table [x=n_points, y=ratio_blind_random_lower_bound_first, col sep=comma] {all_methods_dim_2_eps_0.1_ratio.csv};
    \addlegendentry{Blind random, lower bound first}

    \addplot+ table [x=n_points, y=ratio_blind_greedy, col sep=comma] {all_methods_dim_2_eps_0.1_ratio.csv};
    \addlegendentry{Blind greedy}

    \addplot+ table [x=n_points, y=ratio_greedy, col sep=comma] {all_methods_dim_2_eps_0.1_ratio.csv};
    \addlegendentry{Greedy}

@@

\begin{figure}[ht!]
    \includegraphics[width=0.7\textwidth]{pics/blind-greedy-dims.png}
    \caption{Results of blind greedy spanner for different dimensions.}
    \label{fig:blind_greedy_dimensions}
\end{figure}

@@

    \addplot+ table [x=n_points, y=edges_blind_random, col sep=comma] {all_methods_dim_2_eps_0.1.txt};
    \addlegendentry{Blind random}

    \addplot+ table [x=n_points, y=edges_blind_random_connect_first, col sep=comma] {all_methods_dim_2_eps_0.1.txt};
    \addlegendentry{Blind random, connect first}

    \addplot+ table [x=n_points, y=edges_blind_random_lower_bound_first, col sep=comma] {all_methods_dim_2_eps_0.1.txt};
    \addlegendentry{Blind random, lower bound first}

    \addplot+ table [x=n_points, y=edges_blind_random_connect_first_lower_bound_first, col sep=comma] {all_methods_dim_2_eps_0.1.txt};
    \addlegendentry{Blind random, connect first, lower bound first}

@@

    \addplot+ table [x=Epsilon, y=edges_greedy_dim_2, col sep=comma] {eps_dep.csv};
    \addlegendentry{Greedy, dim = 2}

    \addplot+ table [x=Epsilon, y=edges_blind_greedy_dim_2, col sep=comma] {eps_dep.csv};
    \addlegendentry{Blind greedy, dim = 2}

@@

        \addplot+ table [x=n_points, y=ratio_greedy_dim_0_eps_02_q_1, col sep=comma] {mcgill_perturbed_blind_greedy.csv};
        \addlegendentry{Greedy, $\eps = 0.2$}

        \addplot+ table [x=n_points, y=ratio_blind_greedy_dim_0_eps_02_q_1, col sep=comma] {mcgill_perturbed_blind_greedy.csv};
        \addlegendentry{Blind greedy, $\eps = 0.2$}

        \addplot+ table [x=n_points, y=ratio_greedy_dim_0_eps_05_q_1, col sep=comma] {mcgill_perturbed_blind_greedy.csv};
        \addlegendentry{Greedy, $\eps = 0.5$}

        \addplot+ table [x=n_points, y=ratio_blind_greedy_dim_0_eps_05_q_1, col sep=comma] {mcgill_perturbed_blind_greedy.csv};
        \addlegendentry{Blind greedy, $\eps = 0.5$}

@@

    \addplot+ table [x=n_points, y=sparseness_greedy_dim_0_eps_01_q_1, col sep=comma] {mcgill_perturbed_blind_greedy.csv};
    \addlegendentry{Greedy, $\eps = 0.1$}

    \addplot+ table [x=n_points, y=sparseness_blind_greedy_dim_0_eps_01_q_1, col sep=comma] {mcgill_perturbed_blind_greedy.csv};
    \addlegendentry{Blind greedy, $\eps = 0.1$}

    \addplot+ table [x=n_points, y=sparseness_greedy_dim_0_eps_02_q_1, col sep=comma] {mcgill_perturbed_blind_greedy.csv};
    \addlegendentry{Greedy, $\eps = 0.2$}

    \addplot+ table [x=n_points, y=sparseness_blind_greedy_dim_0_eps_02_q_1, col sep=comma] {mcgill_perturbed_blind_greedy.csv};
    \addlegendentry{Blind greedy, $\eps = 0.2$}

    \addplot+ table [x=n_points, y=sparseness_greedy_dim_0_eps_05_q_1, col sep=comma] {mcgill_perturbed_blind_greedy.csv};
    \addlegendentry{Greedy, $\eps = 0.5$}

    \addplot+ table [x=n_points, y=sparseness_blind_greedy_dim_0_eps_05_q_1, col sep=comma] {mcgill_perturbed_blind_greedy.csv};
    \addlegendentry{Blind greedy, $\eps = 0.5$}

@@

\begin{figure}[hb!]
    \includegraphics[width=0.7\textwidth]{pics/exact_bad_example.png}
    \caption{Example of point set where exact nearest neighbor search cannot be accelerated by maintaining bounds.
    The exact nearest neighbor is the point $p_1$, next point $p_i$ is placed 
    in the curvilinear triange formed by the balls around the query point, $p_2$ and $p_{i-1}$. Even verifying that $p_1$
    is the true nearest neighbor cannot be done without computing all distances $\dist(p_i,q)$. Indeed, every computed
    $\dist(p_i,q)$ allows to exclude the region in the corresponding ball around $p_i$, but all these balls contain only one $p_i$.}
    \label{fig:exact_bad_example}
\end{figure}

@@

\begin{figure}[ht]
    \includegraphics[width=0.7\textwidth]{pics/log_dependency_constant_ratio.png}
    \caption{Ratio $\log(\mbox{computed distances}) / n$ for ANN algorithm. Data is for uniformly distributed
    points.}
    \label{fig:ann_const_ratio}
\end{figure}

\begin{figure}[ht]
    \includegraphics[width=0.7\textwidth]{pics/ann_computed_distances_dimension_dep.png}
    \caption{Number of computed distances for different dimensions. Points are chosen uniformly, $\eps = 0.01$.}
    \label{fig:ann_dimension_dependence}
\end{figure}

>>>>

\title{Metric Spaces with Expensive Distances}
\titlerunning{Metric Spaces with Expensive Distances}
% uncomment for LIPICS 
\author{Anonymous Authors}{Anon}{}{}{}
%\author{Michael Kerber}{TU Graz}{kerber@tugraz.at}{https://orcid.org/0000-0002-8030-9299}{Supported by Austrian Science Fund (FWF) grant number P 29984-N35}
%\author{Arnur Nigmetov}{TU Graz}{nigmetov@tugraz.at}{}{}
%\authorrunning{M. Kerber and A. Nigmetov}
\authorrunning{Anonymous Authors}

@@

%\Copyright{Michael Kerber and Arnur Nigmetov}
\Copyright{Anonymous Authors}
% END uncomment for LIPICS

\def \expDistFigPath {pics/}
\def \expDistDataPath {./}

@@

\begin{figure}[h]
\centering
\includegraphics[width=6cm]{\expDistFigPath  intro_example.eps}
\caption{The compute distances are shown as edges in a graph. Note that the exact distance
of $p_1$ and $p_2$ is unknown. The shortest path from $p_1$ to $p_2$ has length $9$, which clearly
constitutes an upper bound on the distance by triangle inequality.
However, we can also infer that $\dist(p_1,p_2)\geq 3$:
otherwise, the path from $p_3$ to $p_4$ via $p_1$ and $p_2$
would be shorter than the edge $(p_3,p_4)$, again contradicting
triangle inequality.}
\label{fig:1st_example}
\end{figure}

@@

    \addplot+ table [x=n_points, y=edges_quasi_sorted_greedy, col sep=comma] {\expDistDataPath all_methods_dim_2_eps_0.1.txt};
    \addlegendentry{Blind quasi-sorted greedy}

    \addplot+ table [x=n_points, y=edges_quasi_sorted_shaker, col sep=comma] {\expDistDataPath all_methods_dim_2_eps_0.1.txt};
    \addlegendentry{Blind quasi-sorted shaker}

    \addplot+ table [x=n_points, y=edges_wspd, col sep=comma] {\expDistDataPath all_methods_dim_2_eps_0.1.txt};
    \addlegendentry{WSPD}

    \addplot+ table [x=n_points, y=edges_blind_random, col sep=comma] {\expDistDataPath all_methods_dim_2_eps_0.1.txt};
    \addlegendentry{Blind random}

    \addplot+ table [x=n_points, y=edges_blind_random_lower_bound_first, col sep=comma] {\expDistDataPath all_methods_dim_2_eps_0.1.txt};
    \addlegendentry{Blind random, lower bound first}

    \addplot+ table [x=n_points, y=edges_blind_greedy, col sep=comma] {\expDistDataPath all_methods_dim_2_eps_0.1.txt};
    \addlegendentry{Blind greedy}

    \addplot+ table [x=n_points, y=edges_greedy, col sep=comma] {\expDistDataPath all_methods_dim_2_eps_0.1.txt};
    \addlegendentry{Greedy}

@@

    \addplot+ table [x=n_points, y=ratio_blind_random_lower_bound_first, col sep=comma] {\expDistDataPath all_methods_dim_2_eps_0.1_ratio.csv};
    \addlegendentry{Blind random, lower bound first}

    \addplot+ table [x=n_points, y=ratio_blind_greedy, col sep=comma] {\expDistDataPath all_methods_dim_2_eps_0.1_ratio.csv};
    \addlegendentry{Blind greedy}

    \addplot+ table [x=n_points, y=ratio_greedy, col sep=comma] {\expDistDataPath all_methods_dim_2_eps_0.1_ratio.csv};
    \addlegendentry{Greedy}

@@

\begin{figure}[ht!]
    \includegraphics[width=0.7\textwidth]{\expDistFigPath blind-greedy-dims.png}
    \caption{Results of blind greedy spanner for different dimensions.}
    \label{fig:blind_greedy_dimensions}
\end{figure}

@@

    \addplot+ table [x=n_points, y=edges_blind_random, col sep=comma] {\expDistDataPath all_methods_dim_2_eps_0.1.txt};
    \addlegendentry{Blind random}

    \addplot+ table [x=n_points, y=edges_blind_random_connect_first, col sep=comma] {\expDistDataPath all_methods_dim_2_eps_0.1.txt};
    \addlegendentry{Blind random, connect first}

    \addplot+ table [x=n_points, y=edges_blind_random_lower_bound_first, col sep=comma] {\expDistDataPath all_methods_dim_2_eps_0.1.txt};
    \addlegendentry{Blind random, lower bound first}

    \addplot+ table [x=n_points, y=edges_blind_random_connect_first_lower_bound_first, col sep=comma] {\expDistDataPath all_methods_dim_2_eps_0.1.txt};
    \addlegendentry{Blind random, connect first, lower bound first}

@@

    \addplot+ table [x=Epsilon, y=edges_greedy_dim_2, col sep=comma] {\expDistDataPath eps_dep.csv};
    \addlegendentry{Greedy, dim = 2}

    \addplot+ table [x=Epsilon, y=edges_blind_greedy_dim_2, col sep=comma] {\expDistDataPath eps_dep.csv};
    \addlegendentry{Blind greedy, dim = 2}

@@

        \addplot+ table [x=n_points, y=ratio_greedy_dim_0_eps_02_q_1, col sep=comma] {\expDistDataPath mcgill_perturbed_blind_greedy.csv};
        \addlegendentry{Greedy, $\eps = 0.2$}

        \addplot+ table [x=n_points, y=ratio_blind_greedy_dim_0_eps_02_q_1, col sep=comma] {\expDistDataPath mcgill_perturbed_blind_greedy.csv};
        \addlegendentry{Blind greedy, $\eps = 0.2$}

        \addplot+ table [x=n_points, y=ratio_greedy_dim_0_eps_05_q_1, col sep=comma] {\expDistDataPath mcgill_perturbed_blind_greedy.csv};
        \addlegendentry{Greedy, $\eps = 0.5$}

        \addplot+ table [x=n_points, y=ratio_blind_greedy_dim_0_eps_05_q_1, col sep=comma] {\expDistDataPath mcgill_perturbed_blind_greedy.csv};
        \addlegendentry{Blind greedy, $\eps = 0.5$}

@@

    \addplot+ table [x=n_points, y=sparseness_greedy_dim_0_eps_01_q_1, col sep=comma] {\expDistDataPath mcgill_perturbed_blind_greedy.csv};
    \addlegendentry{Greedy, $\eps = 0.1$}

    \addplot+ table [x=n_points, y=sparseness_blind_greedy_dim_0_eps_01_q_1, col sep=comma] {\expDistDataPath mcgill_perturbed_blind_greedy.csv};
    \addlegendentry{Blind greedy, $\eps = 0.1$}

    \addplot+ table [x=n_points, y=sparseness_greedy_dim_0_eps_02_q_1, col sep=comma] {\expDistDataPath mcgill_perturbed_blind_greedy.csv};
    \addlegendentry{Greedy, $\eps = 0.2$}

    \addplot+ table [x=n_points, y=sparseness_blind_greedy_dim_0_eps_02_q_1, col sep=comma] {\expDistDataPath mcgill_perturbed_blind_greedy.csv};
    \addlegendentry{Blind greedy, $\eps = 0.2$}

    \addplot+ table [x=n_points, y=sparseness_greedy_dim_0_eps_05_q_1, col sep=comma] {\expDistDataPath mcgill_perturbed_blind_greedy.csv};
    \addlegendentry{Greedy, $\eps = 0.5$}

    \addplot+ table [x=n_points, y=sparseness_blind_greedy_dim_0_eps_05_q_1, col sep=comma] {\expDistDataPath mcgill_perturbed_blind_greedy.csv};
    \addlegendentry{Blind greedy, $\eps = 0.5$}

@@

\begin{figure}[hb!]
    \includegraphics[width=0.7\textwidth]{\expDistFigPath exact_bad_example.png}
    \caption{Example of point set where exact nearest neighbor search cannot be accelerated by maintaining bounds.
    The exact nearest neighbor is the point $p_1$, next point $p_i$ is placed 
    in the curvilinear triange formed by the balls around the query point, $p_2$ and $p_{i-1}$. Even verifying that $p_1$
    is the true nearest neighbor cannot be done without computing all distances $\dist(p_i,q)$. Indeed, every computed
    $\dist(p_i,q)$ allows to exclude the region in the corresponding ball around $p_i$, but all these balls contain only one $p_i$.}
    \label{fig:exact_bad_example}
\end{figure}

@@

\begin{figure}[ht]
    \includegraphics[width=0.7\textwidth]{\expDistFigPath log_dependency_constant_ratio.png}
    \caption{Ratio $\log(\mbox{computed distances}) / n$ for ANN algorithm. Data is for uniformly distributed
    points.}
    \label{fig:ann_const_ratio}
\end{figure}

\begin{figure}[ht]
    \includegraphics[width=0.7\textwidth]{\expDistFigPath ann_computed_distances_dimension_dep.png}
    \caption{Number of computed distances for different dimensions. Points are chosen uniformly, $\eps = 0.01$.}
    \label{fig:ann_dimension_dependence}
\end{figure}
